Cover letter for the original IJRR submission

### 2017-09-29 reviews_tro/reviewer_response.tex

\quot { (1) A criterion for robust static equilibrium, but it isn't
described until page 12, in Appendix B, Section B. [\dots]  It would really help to describe this MUCH earlier in
the paper (than page 13), if it is a significant novelty of your paper.}

\as{We agree with the remark. To better distinguish between what is novel and not, we modified the paper to move the content of Appendix B.B (now Section VI), as well as D  (the pseudo code of our algorithm, now Section V.B.2) and E (reference to our source code, now Section VII) to the text itself. This results in the criterion being presented much earlier.}\done \jp

\quot{It is a pretty intuitive/typical check for quantifying a
margin for friction conesque violations (using an approximation for
the friction cone, with x and y direction forces considered
independently). }

\as{Thanks for the comment. To our knowledge, what is common in the literature is to reduce the friction coefficient to obtain more conservative bounds on the admissible forces, and thus a ``safety margin'' to the real cone boundary. An issue with this formulation is that the resulting margin may be too small for small forces, and too large for large contact forces. With our formulation, the obtained margin is constant, and thus guaranteed to be greater than the value $b_0$ we compute. We believe this formulation is new, but would be happy to cite any relevant reference if we happen to be wrong. This formulation proves useful to us because the margin $b_0$ allows to choose the more robust candidate for static equilibrium.}\done

@@

\quot{
Provides complete pseudocode.  This is great! However,it's
not fully clear how much of this is novel vs just re-presenting the
same algorithms previously publilshed.	If only code in App. B and C
are new, that should be clear. Conversely, if most of this presentation
was NOT outlined in such detail before, that should also be clear (to
the authors' credit). Either way, a brief clarification in the intro is
recommended.}

\as{We agree with the comment. Indeed, the pseudocode was not present in [20]. We included it in the core of the paper and clarified that it is novel at the end
of Section I.B.}
\qt{Section IV and V of the present paper are thus paraphrasing the original paper, with the exception of Section V.B.2, where we additionally provide the pseudo-code of the algorithm absent from the original paper.}
\done

\quot{(3)	The contribution to provide a solution to P3 doesn't happen
until page 13 (Appendix C) and as with contribution (1), it needs a
better, intuitive summary (briefly) when first mentioned, as well as a
roadmap for the reader as to where it will be described. It is also
ambiguous in the paper whether to what extent (if any) this solution is
novel, versus a presentation of work from [1]. "Provide a solution"
seemed to imply novelty, in the context of this paragraph -- yet later,
the authors (near the end of page 2) state they use a "state-of-the-art
solution to P3", which implies the P3 solution is not a claim of
novelty.}

\as{The main contribution of the paper is the proposal and validation of an algorithm for P1 and P2. 
In the new version of the paper, we do not consider that our solution to P3
is a new contribution, although our formulation differs from [1] (our control variable is the acceleration
of the center of mass, whereas in [1] the control variable is the contact forces). We provide it here for completeness, because it corresponds to our implementation, and 
was required to validate our approach. 
We rephrased the text to clearly indicate that the algorithm is not a contribution. We made the choice not to give an intuition nor describe the solver in the core
of the paper because it is not required to understand the contributions to P1 and P2.}
\qt{ We implemented a solution to $\mathcal{P}_3$ based on a reformulation of our previous work[1]. The implementation of this solution is also provided as an open-source project, which can be integrated directly in our framework. Our solution to $\mathcal{P}_3$ then allows us to verify that the synthesized motions are physically consistent, using our implementation of the state of the art simulator proposed in \textcolor{red}{TODO}.
We do not consider these additions as contributions per se, but describing them here is required to understand the experimental contribution of this paper.}\done

@@

\quot{
(4)	The last contribution is validation using an in-house
simulator, as opposed to more approximate (unrealistic avatar) models
used in previous work.	As with the previous contributions claimed, you
should flesh out just a bit more detail, along with a roadmap to where
a particular contribution appears in the paper, to prepare and guide
the reader a bit more. More precisely, what are the differences between
these two simulation approaches (previous and current)? You need a
better description of what your in-house simulator does and does not
do; a vague citation to another paper based on that simulator [38] is
not sufficient.}

\as{Thanks for the comment. Regarding the absence of details on the dynamic simulator, first of all we apologize because [38] is not the appropriate reference.
We added a section where we describe the simulation in more details.

Regarding the difference between our previous and current work, the postures computed in [20] were never validated in a simulator, nor the complete motions since 
we did not have a solution to P3 at the time. Then the main
difference between the avatars of [20] and the real robots we now address are the joint limitations and auto-collisions probabilities.
The robots are much more constrained in this regard, which makes the planning harder. We thus had to demonstrate that our algorithm is also working
for these models, while maintaining good performance.
We added this missing piece of information to the paper, at the end of Section I.B.}
\textcolor{red}{TODO ANDREA}.

@@

\quot{[\dots] the authors need to better
distinguish what the novel contributions are, and to highlight where
these contributions will be presented in the paper.  Also, many
important details are in the appendices. The paper would be improved
with more intuitive comments/summaries on these details, rather than
just writing (for example), we use a bunch of heuristics, and they are
described in Appendix X.}

\as{We agree. The novel details of the paper have been moved from the appendix to the core of the paper, and are explicitely detailed at the end of Section I.B.} \done

\quot{
Also, where you describe "sensitivity" vs "specificity", it would be
useful to give a more intuitive explanation, example [\dots]}

\as{We updated Section VIII.B.1) to illustrate the meaning of those terms.}
\qt{We then computed the sensitivity and specificity of the reachability condition.  In this context, the sensitivity refers to the percentage of configurations in $C_{Reach}^0$, effectively belonging to \gls{$C_{Contact}^0$}. If a sampled configuration is in $C_{Reach}^0$, but our method is unable to generate a contact configuration from it, as a result the sensitivity
decreases. The sensitivity thus illustrates the confidence we have that any configuration in $C_{Reach}^0$ will effectively lead to a contact configuration.
Conversely, the specificity refers to the percentage of configurations \textbf{not} in $C_{Reach}^0$, effectively \textbf{not} belonging to \gls{$C_{Contact}^0$}.
If a sampled configuration is \textbf{not} in $C_{Reach}^0$, but our method is able to generate a contact configuration from it, as a result the specificity
decreases. The specificity thus illustrates the confidence we have that all configurations that allow contact creation belong to $C_{Reach}^0$ (or informally, the confidence that we are not missing valid solutions).
We thus look for a compromise between sensitivity and specificity.} \done

@@

\quot{
p.8. Need to define ``discretization step'' more precisely, I think(?)… Min distance? Max distance? Nominal distance? In time? In Distance? In all 6 DOF??
}

\as{The distance considered is the distance travelled along the path, computed as a weighted 6D Euclidian variation along it. We updated the text to clarify this}.
\qt{(Section V.B.1) $j$ depends on a user-defined variable, called the discretization step. It corresponds to the ratio between the length of the path $\mathbf{q}^0(t)$, and the number
of configurations selected along it to create the contact configurations. The length of the path is computed as the weighted 6D Euclidian distance
travelled along the it, with a weight of 0.7 for the translation part, and 0.3
for the orientation part.}

\quot{
p.9 ‘addresses highly constrained environment’ – actually, not very well, correct? This was the “exception” case, for which results were rather poor, right? Also a case where “heuristics” are not so good for getting true (reliable) holds on grasp points, etc.?
}

\as{Thanks for the comment. On this particular point however, we wish to maintain our statement. For scenarios such as the car egress, a success rate of $68 \%$ means that the planner
will fail at some point, but more importantly it means that if we restart the planning a few times with the same problem, there is a reasonable probability that it will find a feasible solution.
In this regard, the car egress scenario is effectively addressed, because we found a solution for it. \\
Furthermore, we believe that this result is to be compared with the current state of the art: we are not aware of any other method that was able
to compute a valid motion from scratch, even in simulation. The closest work to this achievement is the one from Escande et al., but the trajectories were manually interpolated
after the contact plan was designed, and even so we do not have any idea on the success rate of the method. \\
Now regarding the heuristics, we guess that the reviewer is referring to the large number of repositionings that occurred in the contact plan. We believe that this issue
is more related to the kinematic and environmental constraints than the heuristic for the choice of contacts. Developing heuristics that would take into account the environment to limit 
contact repositioning is indeed out of the scope of this work.}

@@

\section{Answers to reviewer 4}

\quot{ I suggest that a few papers related to more general motion and contact planning for multi-legged robots can be added. Particularly, the problem of ``foothold selection'' for 4 or 6-legged robots travelling over uneven terrains is a problem related to the one tackled in the paper. It would be nice to select a few representative works (e.g. from the DARPA Learning Locomotion project or
papers related to the ETHZ StarlETH and ANYmal robots) and explain the differences between those approaches and the problem solved in the paper.}
\as{Thanks for the relevant comment. As observed by the reviewer, our state of the art voluntarily focuses on multi-contact planning related work, but it makes sense to cite the suggested papers. We added a paragraph at the end of the state of the art to discuss them. As a matter of fact, we believe these approaches could be considered within
our framework to replace specific parts.}\ndone

\quot{in my opinion a T-RO regular paper should be more self-containing. In several points the authors heavily rely on previous publications, e.g. when presenting the RRT-based planner (Section IV-C). This part should be expanded, describing in more detail what exactly has been changed with respect to the reference paper. }
\as{Thanks for the comment. As it is also suggested by reviewer 1 and the AE, we modified the paper to make it self contained. The references to previous work now solely indicate
that a certain method is not a contribution of the present paper. \\ 
In the particular case of Section IV-C, in fact the code of our implementation of the Bi-RRT with respect to the original one is exactly the same, with a single difference: when
traditionally a collision detection predicate is called to validate a configuration, we replace this predicate with a call to the reachability condition. We modified the text to explain this more clearly. As the Bi-RRT is a famous standard algorithm in the literature, we do not think it is necessary to include it in this paper, but we will gladly do so if required. }
\qt{$C_{Reach}^0$ can be sampled efficiently thanks to Eq.~5, and can thus be used with any standard motion planner.
Our current implementation uses the Bi-RRT planner [25] provided by the HPP software [26].
Our implementation is exactly the same as the pseudo-code of the original planner (which does not detail the configuration validation method). With respect to a ``classic'' implementation, the only difference is that instead of validating a configuration using collision detection, we validate it with the \textit{reachability condition}.} \done

\quot{I am a bit skeptical about the readability of the figures. Some of them seem to be too small (e.g. Fig. 3, Fig. 4), some are of quite poor quality (why the robot silhouettes in Fig. 4 are blurred ?), and the purpose and meaning of Fig. 6 is rather unclear.}
\as{Thanks for the comment, with which we agree in general.  \\We increased the size of Fig. 3 and 4. The robot silhouettes in Fig. 4 are blurred when they describe a previous state of the robots. We added this precision in the text. \\Regarding the quality of our Figures, we are quite surprised by the comment. All of our simulation Figures are taken from high resolution rendering in blender. The other Figures are drawn from manually designed powerpoint vectorial graphics (converted to png). If the reviewer would be so kind as to point out specifically which Figures are of poor quality, we will gadly replace them.\\ As proposed also by reviewer 1, we removed references to previous results obtained in [20], so the Fig. 6 has been removed.}\ndone

\quot{In general, the simulation results presented qualitatively in the paper are a bit hard to interpret without the accompanying video, particularly if someone looks at a b/w print of the paper (as it may be looking in T-RO).}
\as{All of our simulation figures are taken from high resolution rendering in blender, so we wouldn't know how to improve their quality. If the reviewers think the paper would be better without these figures, we are willing to remove them. However, we believe that they would make the paper more dependent on the video, so we would prefer to leave them.}
%\as{\textbf{\textcolor{red}{TODO This is true, although I think he has a crappy printer because I don't think it's that bad. I don't know what to do about it though. I can't simply remove the Figures, because he tends to say he would like the paper to be more independent from the video...}}}\ndone

@@

\quot{ I also found the description of the environment model being
used,
and the way it is used in planning quite superficial. It is said, that
``an octree data structure'' is used to keep the sampled the
end-effector positions, and it is ``intersect'' with the environment.
The idea is OK, I understand that the octree is a sample-based
representation of the effector's workspace, and using some environment
model it is possible to decide if the given configuration can provide a
contact point or not. but how is the environment described ? How
efficient is this intersection operation ? Does this method require the
environment to be described by some specific objects (e.g. polygons) ?
What about the (unavoidable) uncertainty if a world model acquired via
on-line sensing is considered ? Perhaps some of these issues (e.g.
uncertainty) are not in the scope of the paper, but I can imagine that
a robotic researcher who reads the paper and is about to use the open
source software it advertises is very interested in the answers.}

\as{Thanks for the comment. The choice of not detailing too much the environment and octree description goes along with 
our effort to maintain the paper as readable as possible, since it already contains various concepts. However,
we agree that additional information must be added. In Section III we now indicate that the environment is represented as a polygon soup (or 3D mesh), and
that we assume that it is fully known. Although the integration of on-line sensing and state estimation in the planner is an issue currently considered by our team, it is indeed
out of the scope of this paper. We thus also indicate that we assume the environment and state of the robot are fully known. To perform the intersection computation, we use
the fcl collision detection library, which natively proposes the intersection method. We added a reference to the library in the paper. The performance of the method is included
in the ``collision'' column of Table II. We updated the legend of Table II to indicate that this time is negligible in comparison to the other collision detections that occur.
} 
\qt{(Section III) \textbf{The environment} $O$ is defined as the union of the obstacles $O_i$ that it contains. $O$ is represented
as a polygon soup (or mesh), where the normal of each surface is known. No further requirement is needed
by our approach. In this work, we assume the environment is fully known. State uncertainty is out of the scope of the paper. \\
(Legend of Table II)  The Collision column times includes the (negligible) octree intersection operation necessary to retrieve the candidate samples.}
\done

\quot{I appreciate the thorough presentation of results in the paper, however
it seems that this presentation is somewhat biased towards simulation
results. Simulations are perfectly OK for planning algorithms, but if
you decided to present the experimental results (HRP-2), then at least some
illustrations should be included in the paper. now they are only in the
accompanying video.  also a comparison between the results of the same
task(s) executed in dynamic simulation and on the real robot should
convince the reader that the planning method's good performance does
not depend too much on the idealized simulation environment. }
\as{We agree with the remark, and included some illustrations of the experimental results on HRP-2 in the paper. \\
We now elaborate on the validity of the simulation with respect to an experiment on the real robot. As also suggested by reviewer 1, we added a paragraph
in the paper (TODO where) to provide more information on the simulation we use to validate our motions. This simulation is an implementation of the state of the art approaches used by the community. Of course, it does not guarantee that the motion could be executed on the robot, but it is used extensively by the community as a reasonable feasibility test. The suggestion of the reviewer to compare the simulation results with those on the real robot is interesting. However, our robot HRP-2 is not torque controlled, but position controlled, therefore as of today a comparison is not possible.}\ndone

@@

\quot{*I think I understand what the authors mean by solving at
"interactive" rates, but this is worth defining. In my mind, it means
something like "before the human operator gets bored", or roughly > 0.5
Hz. But since the authors make the claim that this is the "first
interactive implementation of a contact planner", then that statement
should be made quantitative and verifiable}
\as{We agree with the relevant comment, and updated the paper accordingly.}\ndone

\quot{* In Section I.A. the authors refer to Bretl's work on multi-contact
planning as the inspiration for the P1,P2,P3 decomposition of the
problem. But it is not, as the authors state, always necessary to
decompose motion planning into choosing a series of quasi-static
configurations. There are a variety of feasible motions with no static
equilibrium configurations, so it is worth mentioning that this
decomposition only solves a subset of motion planning problem.}
\as{We apologize if this was not made clear enough in our first submission. In the conclusion, we indicate that the planner is limited to specific scenarios (where quasi-static motion
is always an option). We updated the state of the art to indicate that the decomposition in general is restricitve.}\ndone

\quot{* Also in Sect. I.A, the authors state that "optimization-based
approaches only converge locally". This is not necessarily true, as
convex or mixed-integer convex optimizations can (sometimes) be solved
to a verifiable global optimum. I would suggest removing this sentence
entirely.}
\as{Thanks for the suggestion. Our phrasing was indeed misleading and inaccurate. Our intention was to explain that the geometry of the scene makes the problem highly non-convex, and
that approaches such as the CIO, if not initialized with an appropriate initial guess, might never get out of the local minima. We followed the reviewer's advice and simply removed the sentence. }\ndone

\quot{* I am not sure why the $C_{reach}$ hypothesis is described as the ``strong''
hypothesis in, e.g., Sect. IV. Is there a corresponding weak
hypothesis?}
\as{We apologize for the confusion, that comes from an inaccurate use of English. What we meant was that the hypothesis is TODO}\ndone

### 2017-09-29 body/appendix_optim.tex
ongoing, added TODO / DONE macros in letter
Review paper according to reviewers suggestions

\begin{equation} \label{eq:lin_prog_rob} \begin{aligned}
%~ \find \quad & \mx{R} \in \Rv{3\times T}, \mxdd{C} \in  \Rv{6\times T} \\
\minimize_{\mathbf{c}(t), \mathbf{\dot{c}}(t), \mathbf{\ddot{c}}(t), \mathbf{w}(t)}  \quad & \sum_{u=0}^{2} \int_{t_u}^{t_u + \Delta t_u}  \ell(\mathbf{c}(t), \mathbf{\dot{c}}(t), \mathbf{\ddot{c}}(t), \mathbf{w}(t)) \text{dt}  \\
\st \quad & \mathbf{\ddot{c}}(t) = dyn(\mathbf{c}(t), \mathbf{\dot{c}}(t), \mathbf{w}(t)) \\
	\quad & \mathbf{A}_u \mathbf{w}(t)  \leq \mathbf{b}_u , \forall t \in  \left[t_u, t_u + \Delta t_u \right[, \forall u \\
	\quad & \mathbf{Y}_u \mathbf{c}(t)  \leq \mathbf{y}_u , \forall t \in  \left[t_u, t_u + \Delta t_u \right[, \forall u \\
	\quad & \mathbf{c}(0)  = \mathbf{c}_{\mathbf{q}_0} \\
	\quad & \mathbf{c}(T)  = \mathbf{c}_{\mathbf{q}_1} \\
	\quad & \mathbf{c}(0)  = \mathbf{\dot{c}}(0)  = \mathbf{\ddot{c}}(0) = \mathbf{0}\\
	\quad & \mathbf{c}(T)  = \mathbf{\dot{c}}(T)  = \mathbf{\ddot{c}}(T) = \mathbf{0}\\
    %~ \quad & ({\vct{r}}\tra\vct{r}=1,  \forall t  ?) \\
    %~ \quad & \delta t \sum\limits_1^{T-1}{\vc{r}^1}\tra {\vc{r}^1} \ge t_{transit} \\
    %~ \quad & \mx{A}_{\int} \mxdd{C} \leq \vc{a}_{\int} \\
    %~ \quad & \mx{V}_{\int} \mxd{C} \leq \vc{v}_{\int} \\
    %~ \quad & \mx{B}_{\int}^{1} \mx{C} \leq \vc{b}_{\int} \\
    %~ \quad & [\vc{c}_0\tra, \vcd{c}_0\tra]\tra = \vc{x}_0 \\
    %~ \quad & \vc{c}_T = \vc{c}_T^{input} \\
\end{aligned} \end{equation}

The cost function $\ell$ is a weighted sum of the angular momentum and center-of-mass acceleration variation over the whole trajectory. 
The system dynamics $dyn$ tie together the position, velocity and acceleration of the center of mass. 
$\mathbf{Y}_u$ and $\mathbf{y}_u$ denote stacked kinematic constraints on the center of mass position, determined by the active contact locations.
The inequalities for each contact are determined in the same way that the reachable workspace is computed in Appendix~\ref{app:rom}, with
the effector serving as root. % (a triangulated object is a polytope, and as such can be described as a set of linear inequalities).

@@

This motion is then validated in our dynamic simulator~\citep{DelPrete2015b}, using the trajectory as a reference postural task for an inverse-dynamics
controller that accounts for all the robot constraints. This final result is shown in the companion video.

>>>>

\begin{equation} \label{eq:qp_rob} \begin{aligned}
%~ \find \quad & \mx{R} \in \Rv{3\times T}, \mxdd{C} \in  \Rv{6\times T} \\
\minimize_{\mathbf{c}(t), \mathbf{\dot{c}}(t), \mathbf{\ddot{c}}(t), \mathbf{w}(t)}  \quad & \sum_{u=0}^{2} \int_{t_u}^{t_u + \Delta t_u}  \ell(\mathbf{c}(t), \mathbf{\dot{c}}(t), \mathbf{\ddot{c}}(t), \mathbf{w}(t)) \text{dt}  \\
\st \quad & \mathbf{\ddot{c}}(t) = dyn(\mathbf{c}(t), \mathbf{\dot{c}}(t), \mathbf{w}(t)) \\
	\quad & \mathbf{A}_u \mathbf{w}(t)  \leq \mathbf{b}_u , \forall t \in  \left[t_u, t_u + \Delta t_u \right[, \forall u \\
	\quad & \mathbf{Y}_u \mathbf{c}(t)  \leq \mathbf{y}_u , \forall t \in  \left[t_u, t_u + \Delta t_u \right[, \forall u \\
	\quad & \mathbf{c}(0)  = \mathbf{c}_{\mathbf{q}_0} \\
	\quad & \mathbf{c}(T)  = \mathbf{c}_{\mathbf{q}_1} \\
	\quad & \mathbf{c}(0)  = \mathbf{\dot{c}}(0)  = \mathbf{\ddot{c}}(0) = \mathbf{0}\\
	\quad & \mathbf{c}(T)  = \mathbf{\dot{c}}(T)  = \mathbf{\ddot{c}}(T) = \mathbf{0}\\
    %~ \quad & ({\vct{r}}\tra\vct{r}=1,  \forall t  ?) \\
    %~ \quad & \delta t \sum\limits_1^{T-1}{\vc{r}^1}\tra {\vc{r}^1} \ge t_{transit} \\
    %~ \quad & \mx{A}_{\int} \mxdd{C} \leq \vc{a}_{\int} \\
    %~ \quad & \mx{V}_{\int} \mxd{C} \leq \vc{v}_{\int} \\
    %~ \quad & \mx{B}_{\int}^{1} \mx{C} \leq \vc{b}_{\int} \\
    %~ \quad & [\vc{c}_0\tra, \vcd{c}_0\tra]\tra = \vc{x}_0 \\
    %~ \quad & \vc{c}_T = \vc{c}_T^{input} \\
\end{aligned} \end{equation}
\adnote{Did you really use w as a variable of your problem? Or do you compute it inside your constraint function as a function of the CoM pos and acc? And what about dL? It appears in the definition of w but then it is no longer there in the problem definition!}
The cost function $\ell$ is a weighted sum of the angular momentum and center-of-mass acceleration variation over the whole trajectory. 
The system dynamics $dyn$ tie together the position, velocity and acceleration of the center of mass. 
$\mathbf{Y}_u$ and $\mathbf{y}_u$ denote stacked kinematic constraints on the center of mass position, determined by the active contact locations.
The inequalities for each contact are determined in the same way that the reachable workspace is computed in Appendix~\ref{app:rom}, with
the effector serving as root. % (a triangulated object is a polytope, and as such can be described as a set of linear inequalities).

@@

%This motion is then validated in our dynamic simulator~\citep{DelPrete2015b}, using the trajectory as a reference postural task for an inverse-dynamics
%controller that accounts for all the robot constraints. This final result is shown in the companion video.

### 2017-09-29 body/conclusion.tex
Review paper according to reviewers suggestions

% !TEX root =  ../main.tex
 \section{Discussion: validity and interest of our contact planner} 
\label{sec:discussion}

@@

\subsection{Addressing $\mathcal{P}_3$}
This paper focuses on contact planning of multi-contact locomotion.It does not address directly the continuous motion generation
. Problem $\mathcal{P}_3$ remains an active research topic~\cite{Carpentier2016, herzog2015trajectory},
for which no off-the shelf solution exists thus far. We thus implemented a dedicated solution, described in Appendix~\ref{app:optim}.
The companion video shows the results that we obtained in simulation and on the real HRP-2 robot.
In these scenarios the robot is able to execute our plans, while respecting the constraints inherent to the problem: joint kinematics and torque limits and
collision avoidance.
%~ \adnote{Probably you should add in appendix an explanation of how you solve $P_3$ and refer to it here.}

>>>>

% !TEX root =  ../main_tro.tex
 \section{Discussion: validity and interest of our contact planner} 
\label{sec:discussion}

@@

\subsection{Addressing $\mathcal{P}_3$}
\adnote{This subsection seems redundant to me}
This paper focuses on contact planning of multi-contact locomotion.It does not address directly the continuous motion generation
. Problem $\mathcal{P}_3$ remains an active research topic~\cite{Carpentier2016, herzog2015trajectory},
for which no off-the shelf solution exists thus far. We thus implemented a dedicated solution, described in Appendix~\ref{app:optim}.
The companion video shows the results that we obtained in simulation and on the real HRP-2 robot.
In these scenarios the robot is able to execute our plans, while respecting the constraints inherent to the problem: joint kinematics and torque limits and
collision avoidance.
%~ \adnote{Probably you should add in appendix an explanation of how you solve $P_3$ and refer to it here.}

### 2017-09-29 body/intro.tex
Review paper according to reviewers suggestions

% !TEX root =  ../main.tex
\newcommand{\Pa}{$\mathcal{P}_1$ }
\newcommand{\Pb}{$\mathcal{P}_2$ }

@@

\noindent \textbf{Comparison with our previous work:}
The present paper is an extension of our conference paper~\cite{tonneauisrr15}. As such, the motion planning algorithm presented here is the same as the one presented in the
previous work. Section~\ref{rbprm} and~\ref{sec:contact} of the present paper are thus paraphrasing the original paper, with the exception of Section~\ref{app:contact}, where we additionally provide the pseudo-code of the algorithm absent from the original paper. We also improve the computational performance of the method by replacing the static equilibrium test with a more efficient test based on an new LP formulation presented in Section~\ref{sec:Equil}.
Compared with the original paper, we also release our source code, and provide the complete procedure to use our planner for any legged robot. We thus provide
all the technical information required to use the planner.

To validate our approach, in this paper we had to implement or integrate other tools based on previous publications from our team. We implemented a solution to $\mathcal{P}_3$ based on a reformulation of our previous work~\cite{Carpentier2016}. The implementation of this solution is also provided as an open-source project, which can be integrated directly in our framework. Our solution to $\mathcal{P}_3$ then allows us to verify that the synthesized motions are physically consistent, using our implementation of the state of the art simulator proposed in TODO.
We do not consider these additions as contributions \textit{per se}, but describing them here is required to understand the experimental contribution of this paper.

>>>>

% !TEX root =  ../main_tro.tex
\newcommand{\Pa}{$\mathcal{P}_1$ }
\newcommand{\Pb}{$\mathcal{P}_2$ }

@@

\noindent \textbf{Comparison with our previous work:}
The present paper is an extension of our conference paper~\cite{tonneauisrr15}. As such, the motion planning algorithm presented here is the same as the one presented in the
previous work. Section~\ref{rbprm} and~\ref{sec:contact} of the present paper are thus paraphrasing the original paper, with the exception of Section~\ref{app:contact}, where we additionally provide the pseudo-code of the algorithm absent from the original paper. We also improve the computational performance of the method by replacing the static equilibrium test with a more efficient test based on a new LP formulation presented in Section~\ref{sec:Equil}.
Compared with the original paper, we also release our source code, and provide the complete procedure to use our planner for any legged robot. We thus provide
all the technical information required to use the planner.

To validate our approach, in this paper we had to implement or integrate other tools based on previous publications from our team. We implemented a solution to $\mathcal{P}_3$ based on a reformulation of our previous work~\cite{Carpentier2016}. The implementation of this solution is also provided as an open-source project, which can be integrated directly in our framework. Our solution to $\mathcal{P}_3$ then allows us to verify that the synthesized motions are physically consistent, using our implementation of a state-of-the-art simulation algorithm~\cite{Kaufman2008}.
We do not consider these additions as contributions \textit{per se}, but describing them here is required to understand the experimental contribution of this paper.
